---
title: "Final Report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In an accelerated program, such as the Master of Data Science (MDS), students are often limited on time. As MDS students, we are highly interested in investigating whether attending office hours helped MDS students complete their lab assignments in a shorter amount of time.

Therefore, the main question addressed by our study is:

How does the number of times a Master of Data Science (MDS) student goes to office hours affect the average number of hours they spend working on labs per week?

To ensure that our study is as causal as possible, we also considered potential confounding variables. Confounding variables investigated in this study include whether someone worked on labs in groups or alone, someone's academic background, the number of years someone has been out-of-school, someone's competency in programming and statistics and whether someone attempted optional questions. 

```{r pressure, echo=FALSE}
```

